---
title: "AI_models"
author: "Angela E. Boag"
date: "June 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aridity index projections for protected areas in US grasslands

This analysis summarizes aridity projections (Seager et al. 2018) across
protected areas in US grasslands. First, we selected three models to represent
different future scenarios for AI in addition to the ensemble model average
projection. 

```{r choosing models}
setwd("~/WestClim/100thMeridian/protected_grasslands")

# Loading mean 2081-2099 AI projections for the entire area covering US plains
# grasslands, and summarizing those projections to identify 3 scenarios. 
# 1: Most positive median AI by 2081-2099; 2: most negative median AI; 
# 3: Lowest mean SD for AI by 2081-2099

models <- read.csv("dAI_2080_2099grasslands.csv")
head(models)

# switch wide to long
library(tidyr)
library(ggplot2)
library(dplyr)
library(plyr)

models2 <- gather(models, modelNo, dAI, dAI_m1:dAI_m18, factor_key = TRUE)

head(models2)
str(models2)

#change all factor variables to numeric
models2$dAI <- as.numeric(models2$dAI)
models2 <- na.omit(models2)

# summary statistics for end-of-century dAI for each model:
ddply(models2, ~ modelNo, summarise, mean = mean(dAI),
      median = median(dAI), sd = sd(dAI),
      min = min(dAI), max = max(dAI)) 

# boxplots
ggplot(models2, aes(x = modelNo, y = dAI)) + 
  geom_boxplot() +
    ylim(c(-8, 8))

```

```{r}
projAI <- read.csv("largestPAs_AI_proj.csv")
head(projAI)

# Create vector of variables that I want to keep in a cleaner version of this 
# dataset. 

vars <- c("FID","Category", "d_Own_Name", "d_Mang_Typ", "d_Mang_Nam", "Unit_Nm",
          "d_State_Nm", "Grassland_Type", "GIS_Acres", "SD_histAI",
          "X2041_60_pos75", "X2081_99_pos75", "X2041_60_neg75", 
          "X2081_99_neg75", "X2041_2060_18model_AI", "X2041_2060_18model_dAI",
          "X2081_2099_18model_AI", "X2081_2099_18model_dAI",
          "X1979_2015_aveAI", "X2041_2060_d18_m18", "X2081_2099_d18_m18", 
          "X2041_2060_dAI_m14", "X2081_2099_dAI_m14", "X2041_2060_dAI_m13",
          "X2081_2099_dAI_m13")
projAI2 <- projAI[vars]
head(projAI2)
          

ggplot(projAI2, aes(x = X1979_2015_aveAI, y = X2081_2099_18model_AI, 
                    shape = Grassland_Type)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  xlim(0, 1.7) +
  ylim(0, 1.7) +
  labs(x="mean annual aridity index 1979-2015", 
       y = "mean annual aridity index 
       (18-model ensemble) 2081-2099") +
   theme_classic()

```



```{r}
# Lollipop plot for average AI of all protected areas
projAI3 <- projAI2[order(X1979_2015_aveAI), ]
          

ggplot(projAI3) +
  geom_segment( aes(x = reorder(Unit_Nm, X2081_2099_18model_AI), xend = Unit_Nm, 
                    y = X1979_2015_aveAI, 
                    yend = X2081_2099_18model_AI), color="grey") +
  geom_point( aes(x = Unit_Nm, y = X1979_2015_aveAI), 
              color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  geom_point( aes(x = Unit_Nm, y = X2081_2099_18model_AI), 
              color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  coord_flip()+
  theme_light() +
  theme(
    legend.position = "none",
    panel.border = element_blank() ) +
  xlab("") +
  ylab("mean annual aridity index (mm)")
  #+ facet_grid(Grassland_Type ~ .)

projAI3[which(Unit_Nm == "Big Stone National Wildlife Refuge"), ]

```

Create new dataframe with state included in Unit Name. 

```{r, graphs for individual grassland types}

# Create new Unit Name variable with the state included. 

projAI3$St_Unit_Nm <- paste(projAI3$d_State_Nm, "-", projAI3$Unit_Nm)
head(projAI3)

```

```{r}
# Lollipop plot for average AI of top 10 dAI protected areas
attach(projAI3)

projAI3_top10end <- projAI3[with(projAI3,order(X2081_2099_18model_dAI)),]

projAI3_top10end <- projAI3_top10end[1:10,]
projAI3_top10end

ggtop10end <- ggplot(projAI3_top10end, 
                     aes(x = St_Unit_Nm, y = X1979_2015_aveAI)) +
  geom_segment( aes(x = reorder(St_Unit_Nm, X2081_2099_18model_AI), 
                    xend = St_Unit_Nm, 
                    y = X1979_2015_aveAI, 
                    yend = X2081_2099_18model_AI), color="grey") +
  geom_point( aes(x = St_Unit_Nm, y = X1979_2015_aveAI), 
              color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  geom_point( aes(x = St_Unit_Nm, y = X2081_2099_18model_AI), 
              color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  coord_flip() +
  theme_classic() +
  theme(
    legend.position = "none",
    panel.border = element_blank() ) +
  xlab("") +
  ylab("mean annual aridity index (mm)") +
  ylim(0, 2.5) +
  theme(text = element_text(size=12, colour = "black")) +
  geom_errorbar(aes(ymin=X1979_2015_aveAI-SD_histAI, 
                    ymax=X1979_2015_aveAI+SD_histAI), 
                width=.2,
                 position=position_dodge(0.05))

ggtop10end
ggsave("ggtop10end.png", dpi = 300, height = 3, width = 7)

```

Lollipop plots for each grassland type. 

```{r}
tall <- projAI3[ which(projAI3$Grassland_Type == 'tall'),]
head(tall)

mixed <- projAI3[ which(projAI3$Grassland_Type == 'mixed'),]
sand <- projAI3[ which(projAI3$Grassland_Type == 'sand'),]
short <- projAI3[ which(projAI3$Grassland_Type == 'short'),]

ggtall <- ggplot(tall, aes(x = St_Unit_Nm, y = X1979_2015_aveAI)) +
  geom_segment( aes(x = reorder(St_Unit_Nm, X2081_2099_18model_AI), 
                    xend = St_Unit_Nm, 
                    y = X1979_2015_aveAI, 
                    yend = X2081_2099_18model_AI), color="grey") +
  geom_point( aes(x = St_Unit_Nm, y = X1979_2015_aveAI), 
              color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  geom_point( aes(x = St_Unit_Nm, y = X2081_2099_18model_AI), 
              color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  coord_flip()+
  theme_classic() +
  theme(
    legend.position = "none",
    panel.border = element_blank() ) +
  xlab("") +
    ylab(" ") +
  ylim(0, 2.5) +
  theme(text = element_text(size=14)) +
  geom_errorbar(aes(ymin=X1979_2015_aveAI-SD_histAI, 
                    ymax=X1979_2015_aveAI+SD_histAI), 
                width=.2,
                 position=position_dodge(0.05))

ggmixed <- ggplot(mixed, aes(x = St_Unit_Nm, y = X1979_2015_aveAI)) +
  geom_segment( aes(x = reorder(St_Unit_Nm, X2081_2099_18model_AI), 
                    xend = St_Unit_Nm, 
                    y = X1979_2015_aveAI, 
                    yend = X2081_2099_18model_AI), color="grey") +
  geom_point( aes(x = St_Unit_Nm, y = X1979_2015_aveAI), 
              color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  geom_point( aes(x = St_Unit_Nm, y = X2081_2099_18model_AI), 
              color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  coord_flip()+
  theme_classic() +
  theme(
    legend.position = "none",
    panel.border = element_blank() ) +
  xlab("") +
    ylab(" ") +
  ylim(0, 2.5) +
  theme(text = element_text(size=14)) +
  geom_errorbar(aes(ymin=X1979_2015_aveAI-SD_histAI, 
                    ymax=X1979_2015_aveAI+SD_histAI), 
                width=.2,
                 position=position_dodge(0.05))

ggsand <- ggplot(sand, aes(x = St_Unit_Nm, y = X1979_2015_aveAI)) +
  geom_segment( aes(x = reorder(St_Unit_Nm, X2081_2099_18model_AI), 
                    xend = St_Unit_Nm, 
                    y = X1979_2015_aveAI, 
                    yend = X2081_2099_18model_AI), color="grey") +
  geom_point( aes(x = St_Unit_Nm, y = X1979_2015_aveAI), 
              color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  geom_point( aes(x = St_Unit_Nm, y = X2081_2099_18model_AI), 
              color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  coord_flip()+
  theme_minimal() +
  theme_classic() +
  theme(
    legend.position = "none",
    panel.border = element_blank() ) +
  xlab("") +
  ylab(" ") +
  ylim(0, 2.5) +
  theme(text = element_text(size=14)) +
  geom_errorbar(aes(ymin=X1979_2015_aveAI-SD_histAI, 
                    ymax=X1979_2015_aveAI+SD_histAI), 
                width=.2,
                 position=position_dodge(0.05))

ggshort <- ggplot(short, aes(x = St_Unit_Nm, y = X1979_2015_aveAI)) +
  geom_segment( aes(x = reorder(St_Unit_Nm, X2081_2099_18model_AI), 
                    xend = St_Unit_Nm, 
                    y = X1979_2015_aveAI, 
                    yend = X2081_2099_18model_AI), color="grey") +
  geom_point( aes(x = St_Unit_Nm, y = X1979_2015_aveAI), 
              color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  geom_point( aes(x = St_Unit_Nm, y = X2081_2099_18model_AI), 
              color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  coord_flip() +
  theme_classic() +
  theme(
    legend.position = "none",
    panel.border = element_blank() ) +
  xlab("") +
  ylab("mean annual aridity index (mm)") +
  ylim(0, 2.5) +
  theme(text = element_text(size=14)) +
  geom_errorbar(aes(ymin=X1979_2015_aveAI-SD_histAI, 
                    ymax=X1979_2015_aveAI+SD_histAI), 
                width=.2,
                 position=position_dodge(0.05))

require(gridExtra)

ggtall

ggall <- grid.arrange(ggtall, ggmixed, ggsand, ggshort, ncol = 1)
g <- arrangeGrob(ggtall, ggmixed, ggsand, ggshort, ncol = 1)
ggsave("aveAI_largePAs.png", dpi = 300, height = 16, width = 8, g)

```

Subset just the 10 protected areas with the highest delta AI. 

```{r}


```